Runtime Feedback in a Meta-Tracing JIT for Efficient Dynamic Languages

### 2011-04-15 talk/icooolps2011/paper.tex
replace one itemize by a sentence, which gives me enough space to resurrect the meta-paragraph.

For constant folding to work, two conditions need to be met:

\begin{itemize}
    \item the arguments of an operation actually need to all be constant,
    i.e. statically known by the optimizer
    \item the operation needs to be \emph{constant-foldable}, i.e. always yield
    the same result given the same arguments.
\end{itemize}

There is one kind of hint for both of these conditions.

@@

It is worth clarifying what a ``constant'' is in this context.  A variable of
the trace is said to be constant if its value is statically known by the
optimizer.

The simplest example of constants are literal values, such as \texttt{1}.
However, the optimizer can statically know the value of a variable even if it
is not a constant in the original source code. For example, consider the
following fragment of RPython code on the left. If the fragment is traced with
\texttt{x} being \texttt{4}, the trace on the right is produced:

@@

The first step in making \texttt{getattr} faster in our object model is to optimize
away the dictionary lookups on the instances. The hints we have introduced in the
two previous sections don't seem to help with the current object model. There is
no trace-elidable function to be seen, and the instance is not a candidate for promotion,
because there tend to be many instances.

@@

XXX separation of fast and slow-changing parts

@@

%\subsection{More General Patterns}
%
%The techniques we used above to make instance and class lookups faster are
%applicable in more general cases than the one we developed them for. A more
%abstract view of maps is that of splitting a data-structure into an immutable part (\eg the map)
%and a part that changes (\eg the storage list). All the computation on the
%immutable part is trace-elidable so that only the manipulation of the quick-changing
%part remains in the trace after optimization.
%
%Similarly, versions can be used to constant-fold arbitrary functions of large data
%structures. The version needs to be updated carefully every time the result of
%this function can change. Therefore this is useful only if the data structure is
%expected to change slowly.

>>>>

For constant folding to work, two conditions need to be met: the arguments of
an operation actually need to all be constant, i.e. statically known by the
optimizer and the operation needs to be \emph{constant-foldable}, i.e. always
yield the same result given the same arguments.
There is one kind of hint for both of these conditions.

@@

It is worth clarifying what a ``constant'' is in this context.  A variable of
the trace is said to be constant if its value is statically known by the
optimizer.
The simplest example of constants are literal values, such as \texttt{1}.
However, the optimizer can statically know the value of a variable even if it
is not a constant in the original source code. For example, consider the
following fragment of RPython code on the left. If the fragment is traced with
\texttt{x} being \texttt{4}, the trace on the right is produced:

@@

The first step in making \texttt{getattr} faster in our object model is to optimize
away the dictionary lookups on the instances. The hints of the previous section
don't seem to help with the current object model. There is
no trace-elidable function to be seen, and the instance is not a candidate for promotion,
because there tend to be many instances.

@@

The technique to make instance lookups faster is applicable in more general
cases. A more abstract view of maps is that of splitting a data-structure into
an immutable part (\eg the map) and a part that changes (\eg the storage list).
All the computation on the immutable part is trace-elidable so that only the
manipulation of the quick-changing part remains in the trace after
optimization.

@@

%\subsection{More General Patterns}
%
%
%Similarly, versions can be used to constant-fold arbitrary functions of large data
%structures. The version needs to be updated carefully every time the result of
%this function can change. Therefore this is useful only if the data structure is
%expected to change slowly.
